---
title: "VA_MentalIllness_Map"
author: "Jenny"
date: "10/10/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

load packages
```{r}
library(jsonlite)
library(tidyverse)
library(ggplot2)
library(choroplethr)
library(choroplethrMaps)
```

Load data from GitHub account https://github.com/mihiriyer/mental
```{r}
#Code from https://github.com/mihiriyer/mental
#load station number - name -visn crosswalk from data dictionary
sta <- xlsx::read.xlsx("NepecPtsdDataDictionary.xlsx", sheetIndex=4, stringsAsFactors=FALSE)

#load le data locally (less resource intensive)
mental <- jsonlite::fromJSON("NEPEC_AnnualDataSheet_MH_FY15.json")

#combine mental and sta
mental <- left_join(mental, sta, by="Station")

#rearrange columns 
mental <- mental[, c(6,3,7,1,2,4,5)]

#remove % and , signs from values and then set to numeric data type (this is needed to sort and perform computations)
mental$Value <- gsub("%", "", mental$Value)
mental$Value <- gsub(",", "", mental$Value)
mental$Value <- as.numeric(mental$Value)
```

Turn multiple variables from character to factor, at the same time.
```{r}
cols <- c("ValueType", "Station", "Station.Name", "Category", "Item")
mental <- mental %>% 
  mutate_at(cols, funs(factor(.)))
```

Create dataframe using Prevalence 

```{r}
View(subset(mental, Category == "Prevalence of Mental Illness"))

Prevalence <- (subset(mental, Category == "Prevalence of Mental Illness"))

Prevalence <- droplevels(Prevalence)
Prevalence <- Prevalence[,-c(4,7)]
Prevalence <- Prevalence %>%
  spread(Item, Value)
## 141 different stations  
```


Import location data for all VA facilities
```{r}
library(readxl)
VA_locations <- read_excel("~/Downloads/mental-master/VA_Locations.xlsx", 
                      sheet = "All VA Facilities")
View(head(VA_locations))
## 1920 locations, most of which are not medical, but administrative or providing non-medical services
```
Make sure zipcodes are read as such
```{r}
library(zipcode)

VA_locations$Zip <- clean.zipcodes(VA_locations$Zip)

names(VA_locations)[1] <- "Station"

```

Most VA locations are administrative or provide non-medical services. Extract only medical locations.
```{r}
library(stringr)

VA_locations_Med <- VA_locations %>%
  filter(str_detect(Facility, 'Medical|Health'))
## 235 locations of possible medical facilities included in Prevalence

VA_locations_Med <- VA_locations_Med[,-c(2,3,6)]
VA_locations_Med <- unique(VA_locations_Med)
## now 219 unique locations
```

Add Zip and State to Prevalence dataframe and calculate totals of diagnoses by State
```{r}
Prevalence <- merge(Prevalence, VA_locations_Med, by = "Station")
Prevalence$State <- as.factor(Prevalence$State)

Prevalence <- Prevalence %>%
  group_by(State) %>%
  mutate(TotStateConfirmedMI = sum(`Number of Service Users with Confirmed Mental Illness`))
```

Add state names in format needed for choropleth map
```{r}
data(state.regions)

Prevalence <- merge(Prevalence, state.regions, by.x = "State", by.y = "abb")

Prevalence_mapdata <- unique(Prevalence[,c(12,13)])
Prevalence_mapdata$value <- Prevalence_mapdata$TotStateConfirmedMI

```
Calculate total number of veterans diagnosed and total as a percentage of all VA patients
```{r}
totalserviceusers <- unique(Prevalence[,c(1,9)])
totalusers <- sum(totalserviceusers$`Total Service Users`)
totaltreatedUS <- sum(Prevalence_mapdata$value)
percentdiagnosed <- totaltreatedUS/totalusers

```

Make title and variables for a reasonably presentable legend
```{r}
title <- "              Over 1,500,000 Veterans (23.6% of all VA patients) Treated for Mental Illness at VA Health Facilities in 2015"

Prevalence_mapdata$factorvalue <- factor(
  cut(Prevalence_mapdata$value, c(3000, 10000, 25000, 50000, 80000, 140000)),
  labels = c("3,000 to 10,000", "10,000 to 25,000", "25,000 to 50,000", "50,000 to 80,000", "80,000 to 140,000"))

Prevalence_mapdata$value_absolute<- Prevalence_mapdata$value
Prevalence_mapdata$value <- Prevalence_mapdata$factorvalue
```

Make map. I don't like how it looks. Did this quickly, but will return to prettify.

```{r}
VA_patientmap <- state_choropleth(Prevalence_mapdata, title = title,
                                  legend = "Veterans Treated Per State") +
                                  labs(caption="@jblistman. Data source: Veterans Administration NEPEC Annual Data Sheet 2015")

VA_patientmap
```
